Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-10-26 doc/paper-lcr.tex
added new figures
1st draft complete

\begin{abstract}

Strong gravitational lenses can provide high precision cosmological distance
measurements; one of the most important sources of systematic error in these
estimates is the additional mass along the line of sight to the source. Weak
lensing by galaxies and groups in this beam can contribute an ``external
convergence'' $\xkappa$ that dominates the  uncertainty in the inferred
distances; likewise, it may also contaminate estimates of the luminosity of
objects at high redshift (such as reionisation-era proto-galaxies, and type Ia
supernovae standard candles).  We characterise this uncertainty by marginalising
over a probability density function for $\xkappa$: here, we investigate the use
of a simple halo model to estimate $\Pr(\xkappa)$ given noisy estimates of the
photometric redshift and stellar mass of galaxies observed in optical imaging
surveys of the field in question. We use mock catalogs from the Millenium
Simulation and compare our reconstructed $\Pr(\xkappa)$ to ray-traced $\kappa$ values,
we find that our model produces an unbiased estimator of $\kappax$,
For each lightcone the individual $\Pr(\xkappa)$ from this reconstruction process are
typically 1.8 times narrower than the global distribution of $\xkappa$ given perfect knowledge 
of the halo mass and redshift. After incorporating uncertainties on redshift and halo and stellar masses
we typically find that the reconstruction yields an improvement factor of 1.5. For each line of sight the
 $\Pr(\xkappa)$ can be formed before the investment of light-curve monitoring. By following up only the 
most costrained lines of sight a factor of 2 improvement over the global $\pr(\kappax)$ should be 
achievable.
\end{abstract}

@@

Averaged over large areas of sky, the weak gravitational lensing effect can be
used to constrain the statistical properties of the matter density in the
Universe: every distant object we observe has had its shape distorted, and size
and total brightness magnified (or demagnified) by a compound lens constructed
from all the mass distributed between us and it.

This fact makes weak gravitational lensing a potentially important source of
systematic error for any estimate of luminosity or distance, an issue
raised for \eg 
type Ia supernovae by \citet[][]{Holz+Wald1998,Linder+Holz2004}, for
gamma ray bursts by \citet[][]{Oguri+Takahashi2006,Wang+Dai2011}, 
and for high redshift galaxies by \citet{BradacEtal2009}, among others. 
Along lines of sight containing strong gravitational lenses,
the effect has been found to be
particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios \citep[][]{DalalEtal2005}.
In time delay lens cosmography,
\citet{SuyuEtal2010} found that  the lensing effect of mass structures along the
(unusually over-dense)  line of sight to the quadruply-imaged radio source
B1608$+$656, the so-called ``external convergence,'' was the dominant source
of uncertainty in their 6\% measurement of Hubble's constant.

How can we correct for this weak lensing error, and make
accurate measurements through the Universe? When inferring global quantities
(such as the cosmological parameters), averaging the results from a large
number of independent objects will tend to reduce the error, as the
ensemble-averaged convergence approaches zero for an unbiased sample
\citep[\eg][]{Linder+Holz2004}. If the sample is biased, however, some
residual systematic error will remain while the statistical uncertainty
decreases -- and this will be the case for any sample selected by the
brightness of the lensed object, or any other quantity that depends on
$\xkappa$. The latter class includes time delay lenses, which could be
either detected or selected 
based on their image separations, magnifications or quadruple image
configurations, all of which either increase or become more likely with
increasing $\kappa$. Large
samples of lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010} -- but these
samples will be image separation-limited, due to the surveys' relatively poor
($\sim 1$~arcsec) image quality. Some of the current sample of known time
delay lenses were detected in surveys like this: we might expect them to be
somewhat biased towards high values of $\xkappa$.

There are two broad approaches one could take at this point. First, one could
seek to characterise the lensed objects' selection function, and use this
understanding to derive a suitable correction to cosmological parameters
inferred assuming no $\xkappa$ at all. As pointed out by  \citet[][among
others]{Oguri+Takahashi2006}, to do this one would need  information about the
unlensed population of objects, and the population of lenses, and how the
objects were detected -- or to infer these hierarchically during the analysis.
Alternatively, one could simply accept all lenses provided, regardless of
their selection, and try to use any additional information that is available
to estimate the external convergence in each system on a case by case basis,
and hence optimize their joint analysis to produce a better-constrained result
\citep[as suggested by][]{Keeton+Zabludoff2004}. In this work we follow this
second approach, and use measurements of the positions, magnitude and colours
of galaxies close to lens lines of sight to try to infer the
external convergence at the lens position.

Attempts to date at reconstructing the local and  line of sight mass
distribution in strong lens fields (what \citeauthor{WongEtal2011} call ``the
full environment'') have focused on photometric and  spectroscopic surveys to
detect galaxy groups and ascertain galaxies' membership,  and on understanding
the external shear apparently required by the strong lens models. Surveys by
\citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies both hosting and lying  along the line of sight
to known gravitational lenses. Initial estimates of the effect of these
structures suggested the importance of assigning mass to both galaxy and group
halos, and to allowing for significant uncertainty in the centroids of group
halos. The probability density function (PDF) for the external convergence for
that lens, $\Pr(\xkappa)$, is the target for any reconstruction analysis, and
is the key quantity for any ensemble analysis, since by marginalising over it
we can correctly propagate the uncertainties due to weak lensing. Coming
closest to estimating this quantity from survey data, \citet{WongEtal2011} ran
Monte Carlo reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line of
sight, and compared the resulting predicted shear distributions with the
external shear demanded by the strong lens model. Their simulations show most
of the shear being generated by bright galaxies within 2~arcminutes of the
lens, and that both line of sight structures and the group of galaxies in each
lens plane contribute significant proportions of the shear. They find
significant discrepancies between the lens model shear and the Monte Carlo
predictions, and are unable to distinguish between the environment modelling
and the strong lens modelling as the cause of these discrepancies.

These observational investigations have provided very useful insight into the
problem.  Coming from a more theoretical direction, large numerical
simulations have been used to predict, from first principles, the distribution
of likely external shears and convergences along strong lens lines of sight. 
\citet{Holder+Schechter2003} and \citet{Dalal+Watson2004} carried out
ray-tracing calculations in N-body simulations to estimate the  distribution
of external shear values.  \citet{HilbertEtal2009} performed similar ray-tracing
experiments through the Millenium Simulation \citep{SpringelEtal2005},
generating a predicted $\xkappa$ at every position in a simulated sky. The
first step towards using this excellent calibration resource was taken by
\citet{SuyuEtal2010}, who selected Millenium Simulation lines of sight that a)
were found to contain a strong lens and b) had comparable apparent galaxy
overdensity in a 45 arcsec radius aperture down to $i < 24$. This overdensity
could be compared to the overdensity of the B1608$+$656 line of sight
\citep{FassnachtEtal2011}, and hence an estimate of  $\Pr(\xkappa)$ could be
made.

In this paper, we bring these observational and theoretical approaches
together and  use numerical simulations to {\it test and calibrate} a
reconstruction of the  external convergence. Ray-tracing through the
simulations gives a ``truth'' that we can attempt to recover.  In this paper
we use the Millenium Simulation mock galaxy catalogs and ray-traced lensing
effects to test the accuracy of line of sight mass reconstructions.
Probabilistically ssigning mass and redshift to every observed galaxy in the
field around a time delay lens, we generate Monte Carlo sample line of sight
mass distributions, and estimate the external convergence due to each one.
This gives us a $\Pr(\xkappa)$ that can be propagated into further joint
inferences. We focus on time delay lens cosmography, and quantify our results
in terms of the uncertainty on the mean time delay distance to a test sample
of $N$ lenses (where for clarity each lens has the same deflector and source
redshift, and the same lens model and time delay uncertainty).

We aim to answer the following questions:
\begin{itemize}
\item Faint galaxies and other dark structures will not appear in a photometric
object catalog, but they will contribute convergence at some level. How much of
the total external convergence in a time delay lens system comes from visible
galaxies? How does this change as a function of magnitude cut? 
\item Can the true (ray-traced) convergence be recovered by halo model
reconstruction? What scatter and residual bias are induced in the ensemble 
distance estimate? Which aspects of the model dominate these uncertainties? 
\item When faced with a newly-detected lens, surrounded by galaxies on the sky,
we have some choices to make when planning follow-up observations.
Which are the most important galaxies, with regard to the external convergence
produced? Can nearby groups and clusters be straightforwardly accounted for? If
lenses with such massive structures nearby are discarded, what impact does that
have on the distance accuracy?
\end{itemize}

This paper is organized as follows. In~\Sref{sec:theory} we review the relevant
gravitational lensing theory, before introducing our simple reconstruction 
model in \Sref{sec:model}.  We then test this model in two phases: first, in
\Sref{sec:knownMh+z}, with known redshift and halo mass for every galaxy in a
lens field, in order to quantify the irreducible uncertainty due to unseen mass,
and second, in \Sref{sec:obsMstar+z}, with realistic observational uncertainties
on the field galaxies' stellar masses and redshifts. \comments{We then propagate all these
uncertainties through to the time delay distance in a toy sample of 100 lenses,
in \Sref{sec:distances}.} We discuss our results in \Sref{sec:discuss} before
concluding in \Sref{sec:conclude}.

@@

@@

%Time delay distances, and the mass sheet degeneracy - how kappa is degenerate in
%strong lens models (and magnifications).

@@

Strong lensing occurs when a massive object and background source are almost perfectly
aligned along a line of sight. The light from the background source is deflected by the
lens galaxy; this deflection allows multiple images of the background source to form
at stationary points of the time delay function. For an isolated lens, the time delay
function can be calculated from
\be \label{eq:T} 
\Delta t(\bmath{\theta},\bmath{\beta}) = \frac {1}{c} \frac{D_{\rm od} D_{\rm os}}{D_{\rm ds}} (1+z_{\rm d})\, \phi(\bmath{\theta},\bmath{\beta}),
\ee
where $\bmath{\theta}$ is the observed source position, $\bmath{\beta}$ is the 
unlensed source position, $z_{\rm d}$ is the redshift of the lens, $\phi(\bmath{\theta},\bmath{\beta})$ is
the Fermat potential. The Fermat potential is given by
\be \label{eq:FP}
\phi(\bmath{\theta},\bmath{\beta})\equiv \left[\frac{(\bmath{\theta}-\bmath{\beta})^2}{2}-\psi(\bmath{\theta}) \right], 
\ee
where $\psi(\bmath{\theta})$ is the lens potential, derived from the projected dimensionless
surface mass density, $\kappa(\bmath{\theta})$, by 
\be \label{eq:psikappa}
\kappa(\bmath{\theta})=\frac{1}{2}\nabla^2\psi(\bmath{\theta}).
\ee

Strong lenses sit in our real Universe -- they are not truely isolated;
there is some mass along the line of sight coming from intervening structures.
These structures may be physically associated with the lens e.g. other members
of a group or cluster, or they may be the result of chance alignments at any redshift. 
Whilst it is rare for three galaxies to line up well enough for both of the background sources
to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, the shear scale of dark matter halos
makes a partial alignment -- with the additional structure acting as a weak lens -- 
a near certainty, indeed \citet{Vale+White2003} and \citet{HilbertEtal2007} find that there
are no empty lightcones in a realistic universe. These additional structures induce an external shear
which can be recovered in the mass-modelling of the lens, but they can also introduce a convergence that
is undetectable from the image positions, shapes and relative fluxes; {\it the
mass sheet-degeneracy} \citep{FalcoEtal1985}.

If the intrinsic source brightness changes, the observed brightness in each image does not change
simultaneously - there is a time delay, $\Delta t$, between the brightness change in
each image. Despite the invariance of the image positions and fluxes, under the addition
of an external convergence (and appropriate re-scaling of the lens potential) the relative Fermat potential between the images is {\it not}
invariant, and so the time delay changes. As such it is necessary to include $\kappax$
in the lens modelling, if cosmological parameters are to be estimated accurately and
precisely from observed time delays \citep{SuyuEtal2010}.

\comments{
If the mass sheet is not physically associated with the lens galaxy, it can lie at any redshift. The effects of such a mass-sheet can be projected back onto the lens plane in the form of an effective convergence. We define this effective convergence in two limits.
\begin{itemize}
\item No significant perturbations along the line of sight, and \\
\item A single dominant perturber, relative to which all other perturbations are small.
\end{itemize}
We will refer to these cases as the magnification case (subscript $\mu$ in formulae), and the strong lensing case (subscript SL in formulae), respectively. We will ignore the additional case, of a compound lens, with several significant perturbers.

The magnification case has no primary lens plane, and is chiefly relevant for the magnifications of distant objects. In the absence of a primary lens plane, we are free to define the effective convergence on any point; the effective convergence is the same, independant of the plane choosen. It can be shown \citep[\eg][]{xxx} that if there are several mass sheets at redshifts $\{z_{p}\}$, with convergences $\{\kappa_p =\Sigma_{p}/\Sigma_{\mathrm{cr}}(z_p)\}$, the effective convergence is given by
\be \label{eq:kappasummu}
\kappa_{\mathrm{effective,\mu}}=\sum_{p}\kappa_p.
\ee

In the stong lensing case, the presence of a significant perturber alters the
geometry of the system, and provides a primary lens plane. It is the primary
lens plane upon which we need to know $\kappa_{\mathrm{effective}}$, since
this is where the mass-sheet degeneracy is defined. Following Equations 20--32
of \citet{keeton2003}\flag{keeton2003}, it can be shown that the relevant
effective convergence is given by 
\be \label{eq:kappasumSL}
\kappa_{\mathrm{effective,SL}}=
{  
(1-\beta_{12})(\kappa-\beta_{12}(\kappa^2-\gamma^2))
\over
(1-\beta_{12}\kappa)^2-(\beta_{12}\gamma)^2
}
\ee
where $\kappa$ and $\gamma$ are, respectively, the convergence and shear
caused by the perturber, in their own plane and $\beta_{12}$ is a geometrical
scaling factor defined by
\be\label{eq:beta}
\beta_{12}={
D_{12} D_{\mathrm{os}}
\over
D_{\mathrm{o}1} D_{1\mathrm{s}}
}
\ee
where the D's are again angular diameter distances: the 1 (2) in the subscript
refers to the lower (higher) redshift object of the lens and perturber. For
the case that the mass sheet is in the lens plane, $\beta_{12} = 0$ and
$\kappa_{\mathrm{eff}}=\kappa$; if the mass-sheet is in the plane of the
observer, or source, $\beta_{12} = 1$ and $\kappa_{\mathrm{eff}}=0$; for a
mass-sheet at any other redshift $0<\beta_{12}<1$. In the case of several mass
sheets the effective convergences sum. In the limit that both $\kappa$ and
$\gamma$ are small, Equation \ref{eq:kappasumSL} reduces to 
\be \label{eq:kappasumSLweak}
\kappa_{\mathrm{effective,SL}}=(1-\beta_{12})\kappa;
\ee
it is this $\kappa_{\mathrm{effective,SL}}$ that we use as $\kappax$ for the
strong lensing time delay analysis in \ref{sec:Dt}.
}

\tom{The all-important calibration of the reconstruction has to be done
against Stefan's ray-traced kappas though -- so it is these that we have to
predict PDFs for, first and foremost. THEN we have to consider the question of
which kappa to use to correct the time delay distances: Stefan's kappa (as in
B1608 or RXJ1131), or kappa-keeton.}

The time-delay distance is defined as
\be \label{eq:dt}
\Dt \equiv (1+\zd) D_{\rm d} D_{\rm s}/D_{\rm ds},
\ee
with which Equation \ref{eq:T} can be rephrased as
\be
\Delta t(\bmath{\theta},\bmath{\beta})  =  \frac{\Dt}{c} \, \phi(\bmath{\theta},\bmath{\beta})    ,
\ee 
Since the time delays are proportional to $1/ (1- \kappax)$, if there is any external convergence, that is not included in the lens
modeling, then the time delay distance -- inferred assuming $\kappax = 0$ -- will be
$(1-\kappax)$ less than the true value of $\Dt$.
\be 
\label{eq:MassSheet:H0bias}
\Dt^{\rm{true}}=(1-\kappax) \Dt^{{\kappax = 0}}.
\ee
We can overcome this degeneracy if we have additional knowledge of $\kappax$.

For mass-sheets that are physically associated with the lens galaxy, the mass sheet
 affects the stellar dynamics of the
lens galaxy, so dynamical observations can break the internal
mass-sheet degeneracy by providing an additional estimate of the lens galaxy's mass
\citep[e.g.,][]{xxx}.

In this work our aim is to quantify the lensing effect of mass-sheets that are not physically associated with the lens galaxy -- mass-sheets
lying infront or behind the lens -- and the uncertainties induced by reconstructing
those mass-sheets using simple prescriptions applied to potential astronomical observables.
\comments{We also aim to quantify how uncertainties in the
estimations of $\kappax$ propogate into uncertainties on the cosmological parametres
which can be inferred from measurements of $\Dt$.}

@@

\section{Estimating external convergence: The halo model approximation}
\label{sec:model}

%Weak lensing by NFW halos. Accounting for mass along the line of sight.
%Approximations to full ray tracing:
%
%Halos. N-body sims provide mass framework. MS halo catalog. Truncated spherical
%NFW approx based on virial masses. Stellar mass. Halo map corresponding to
%ray-traced kappa map, for illustration. Small-scale structure

For real lines of sight, we would like to estimate the external convergence as best we can given all
data available. Typically this data is in the form of a photometric galaxy
catalog, containing positions, brightnesses, colours, and possibly spectra for each galaxy in the
field of the lens, down to some magnitude
limit. Since we can not observe the surface mass density of each halo
directy, we need some way of estimating
the mass of each halo in the catalog, so that we can compute its
contribution to the total $\kappax$ along the line of sight. This mass assignment recipe will be uncertain, but
can be calibrated with cosmological simulations.

% A source with known unlensed brightness e.g. a strongly lensed type Ia
% Supernova \flag{cite some SNe papers} or a source of known unlensed size \flag{ideas?}
% would break the mass-sheet degeneracy and provide a value for $\kappax$, but
% \flag{no?} such systems are currently known.

\subsection{Ray-tracing through simulated Lines of Sight}
\label{subsec:raytracing}
Cosmological dark matter simulations can be used to generate mock lines of
sight and can be raytraced through to give each mock sightline a $\kappax$ value...
\todo{Stefan}{Write section on how the ray tracing was done}

\subsection{Generating mock lightcones from simulations}

\todo{Stefan}{Write section on how halos are extracted from the MS n-body particles}

Transforming the photometometric galaxy catalogue into a  catalogue of
halo masses, positions, and redshifts, will enable us to attempt a
reconstruction of the convergence induced by every halo near a line of sight.
This is the subject of \ref{subsec:halos}. We also need to account for
the mass in the Universe that is not associated with the galaxies we can see
-- this is the subject of \ref{subsec:voids}.

>>>>

\begin{abstract} 
Strong gravitational lenses can provide high precision cosmological distance
measurements; one of the most important sources of systematic error in these
estimates is the additional mass along the line of sight to the source. Weak
lensing perturbations by galaxies and groups in this beam can contribute an ``external
convergence'' $\kappa$ that dominates the  uncertainty in the inferred
distances; likewise, it may also contaminate estimates of the luminosity of
objects at high redshift (such as reionisation-era proto-galaxies, and type Ia
supernovae standard candles).  We characterise this uncertainty by marginalising
over a probability density function for $\kappa$: here, we investigate the use
of a simple halo model to estimate $\Pr(\kappa)$ given noisy estimates of the
photometric redshift and stellar mass of galaxies observed in optical imaging
surveys of the field in question. We use mock catalogs from the Millenium
Simulation and compare our reconstructed $\Pr(\kappa)$ to ray-traced $\kappa$ values,
we find that our model produces a precise estimator of $\kappa$ with only small biases.
For each lightcone the individual $\Pr(\kappa)$ from this reconstruction process are
typically 1.8 times more precise than the global distribution of $\kappa$ given perfect knowledge 
of the halo mass and redshift. After incorporating uncertainties on redshift and halo and stellar masses
we typically find that the reconstruction yields an improvement factor of 1.4. For each line of sight the
 $\Pr(\kappa)$ can be formed before the investment of light-curve monitoring: by following up only the 
most costrained lines of sight a factor of 2 improvement over the global $\pr(\kappax)$ should be 
achievable.
\end{abstract}

@@

Every distant object we observe has had its shape distorted, and size
and total brightness magnified (or demagnified) by a compound lens constructed
from all the mass distributed between us and it. As \citet{Vale+White2003} and \citet{HilbertEtal2007}
showed, there are no empty lightcones in our universe. This fact makes gravitational lensing a potentially important source of systematic error for any estimate of luminosity or distance, an issue
raised for \eg type Ia supernovae by \citet[][]{Holz+Wald1998,Linder+Holz2004}, for
gamma ray bursts by \citet[][]{Oguri+Takahashi2006,Wang+Dai2011}, 
and for high redshift galaxies by \citet{BradacEtal2009}, among others.

Along lines of sight containing strong gravitational lenses,
the effect has been found to be
particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios \citep[][]{DalalEtal2005}.
\citet{SuyuEtal2010} found that in time delay lens cosmography the
lensing effect of mass structures along the (unusually over-dense) line of
sight to the quadruply-imaged radio source B1608$+$656, the so-called ``external
convergence'' $\kappax$, was the dominant source of uncertainty in 
their 6\% measurement of Hubble's constant.

The problem is that one simply doesn't know
how magnified and distorted the source has been because the unlensed
source is unobservable. The external shear can potentially be recovered
in the mass-modelling of the lens, but the convergence is undetectable
from the image positions, shapes and relative fluxes; this is the
famous {\emph{  mass sheet-degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

How can we correct for this perturbative lensing and make
accurate measurements through the Universe? When inferring global quantities
(such as the cosmological parameters), averaging the results from a large
number of independent objects will tend to reduce the uncertainty, as the
ensemble-averaged convergence approaches zero for an unbiased sample
\citep[\eg][]{Linder+Holz2004}. If the sample is biased, however, some
residual systematic error will remain while the statistical uncertainty
decreases -- and this will be the case for any sample selected by the
brightness of the lensed object, or any other quantity that depends on
$\kappa$. Large
samples of lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010} -- but these
samples will be image separation-limited, due to the surveys' relatively poor
($\sim 1$~arcsec) image quality. There is a possibility that
they are somewhat biased towards high values of $\kappa$, although 
\citet{Fassnacht+Lubin2002} \todo{Anyone}{check this is the right Fassnacht paper} found no evidence of lenses lying along over-dense lines of sight.

To date, attempts at reconstructing the local and line of sight mass
distribution in strong lens fields (what \citeauthor{WongEtal2011} call ``the
full environment'') have focused on photometric and  spectroscopic surveys to
detect galaxy groups and ascertain galaxies' membership in order to understand
the external shear apparently required by the strong lens models. Surveys by
\citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies both hosting and lying  along the line of sight
to known gravitational lenses. Initial estimates of the effect of these
structures suggested the importance of assigning mass to both galaxy and group
halos, and to allowing for significant uncertainty in the centroids of group
halos. The probability density function (PDF) for the external convergence for
that lens, $\Pr(\xkappa)$, is the target for any reconstruction analysis, and
is the key quantity for any ensemble analysis, since by marginalising over it
we can correctly propagate the uncertainties due to line-of-sight perturbations. Coming
closest to estimating this quantity from survey data, \citet{WongEtal2011} ran
Monte Carlo reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line of
sight, and compared the resulting predicted shear distributions with the
external shear demanded by the strong lens model. Their reconstructions show most
of the shear being generated by bright galaxies within 2~arcminutes of the
lens, and that both line of sight structures and the group of galaxies in each
lens plane contribute significant proportions of the shear. They find
significant discrepancies between the lens model shear and the Monte Carlo
predictions, and are unable to distinguish between the environment modelling
and the strong lens modelling as the cause of these discrepancies.

These observational investigations have provided very useful insight into the
problem.  Coming from a more theoretical direction, large numerical
simulations have been used to predict, from first principles, the distribution
of likely external shears and convergences along strong lens lines of sight. 
\citet{Holder+Schechter2003} and \citet{Dalal+Watson2004}\todo{phil}{Dalal+Watson2004 doesn't seem to exist} carried out
ray-tracing calculations in N-body simulations to estimate the  distribution
of external shear values.  \citet{HilbertEtal2009} performed similar ray-tracing
experiments through the Millenium Simulation \citep{SpringelEtal2005},
generating a predicted $\kappa$ at every position in a simulated sky. \citet{HilbertEtal2009}
found that \MS lines of sight with strong lenses were not biased towards high $\kappa$,
although selection functions for discovering a strong lens were not taken into account.
The first step towards using this excellent calibration resource was taken by
\citet{SuyuEtal2010}, who selected Millenium Simulation lines of sight that
 had comparable apparent galaxy
overdensity in a 45 arcsec radius aperture down to $i < 24.5$. This overdensity
could be compared to the overdensity of the B1608$+$656 line of sight
\citep{FassnachtEtal2011}, and hence an estimate of  $\Pr(\kappa)$ could be
made. B1608$+$656 is indeed a high $\kappa$ line of sight, but it does not reside
in a typical lens environment.

In this paper, we bring these observational and theoretical approaches
together and  use numerical simulations to {\it test and calibrate} a
reconstruction of the  external convergence. Ray-tracing through the
simulations gives a ``truth'' that we can attempt to recover.  In this paper
we use the Millenium Simulation mock galaxy catalogs and ray-traced lensing
effects to test the accuracy of line of sight mass reconstructions.
Probabilistically assigning mass and redshift to every observed galaxy in the
field around a time delay lens, we generate Monte Carlo sample line of sight
mass distributions, and estimate the external convergence due to each one.
By using the Millenium siumlation as a calibration tool, we can convert the
total contribution from halos, $\Pr(\kappah)$, into a PDF for the likely convergence
along the line of sight $\Pr(\kappa)$.

We aim to answer the following questions:
\begin{itemize}
\item Faint galaxies, filaments and dark structures will not appear in a photometric
object catalog, but they will contribute convergence at some level. How much of
the total external convergence in a time delay lens system comes from visible
galaxies? How does this change as a function of magnitude cut? 
\item Can the true (ray-traced) convergence be recovered from a calibrated halo model
reconstruction? What scatter and residual bias are induced by the reconstruction
process? Which aspects of the model dominate these uncertainties? 
\item When faced with a newly-detected lens, surrounded by galaxies on the sky,
we have some choices to make when planning follow-up observations.
Which are the most important galaxies, with regard to the external convergence
produced? Can nearby groups and clusters be straightforwardly accounted for? If
lenses with such massive structures nearby are discarded, what impact does that
have on the distance accuracy?
\end{itemize}

This paper is organized as follows. In~\Sref{sec:theory} we review the relevant
gravitational lensing theory, before introducing our simple reconstruction 
model in \Sref{sec:model}.  We then test this model in two phases: first, in
\Sref{sec:knownMh+z}, with known redshift and halo mass for every galaxy in a
lens field, in order to quantify the irreducible uncertainty due to unseen mass,
and second, in \Sref{sec:obsMstar+z}, with realistic observational uncertainties
on the field galaxies' stellar masses and redshifts. In \Sref{sec:biases} we
investigate potential systematic errors induced by selecting a subset of sightlines.
We discuss our results in \Sref{sec:discuss} before
concluding in \Sref{sec:conclude}.

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{4 different views of a lightcone around a typical line of sight. This line of sight is slightly overdense with a $\kappa$ of 0.03. Top left show the positions of galaxies, the area of the circles is proportional to observed $i$-band flux; the brightest object shown has an $i$ magnitude of 17.7. Top Centre show the angular sizes of halos on the sky; red regions are within the NFW scale radius of a halo and blue shows the virial radius of each halo - there are clearly no empty lightcones. Top right shows the individual $\kappa$ contributions of each halo, assuming the \citet{BMO} profile and the \citet{Neto2007} mass-concentration relation. Bottom a view along the redshift axis, showing only halos with $|x|<0.3$ arcmin; the area of the points is scaled by halo mass; the most massive halo shown has $1.6\times10^{12}\Msun$; the optical axis is the dashed line whilst the dotted lines are the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

@@

\comments{
Strong lensing occurs when a massive object and background source are almost perfectly
aligned along a line of sight. The light from the background source is deflected by the
lens galaxy; this deflection allows multiple images of the background source to form
at stationary points of the time delay function. For an isolated lens, the time delay
function can be calculated from
\be \label{eq:T} 
\Delta t(\bmath{\theta},\bmath{\beta}) = \frac {1}{c} \frac{D_{\rm od} D_{\rm os}}{D_{\rm ds}} (1+z_{\rm d})\, \phi(\bmath{\theta},\bmath{\beta}),
\ee
where $\bmath{\theta}$ is the observed source position, $\bmath{\beta}$ is the 
unlensed source position, $z_{\rm d}$ is the redshift of the lens, $\phi(\bmath{\theta},\bmath{\beta})$ is
the Fermat potential. The Fermat potential is given by
\be \label{eq:FP}
\phi(\bmath{\theta},\bmath{\beta})\equiv \left[\frac{(\bmath{\theta}-\bmath{\beta})^2}{2}-\psi(\bmath{\theta}) \right], 
\ee
where $\psi(\bmath{\theta})$ is the lens potential, derived from the projected dimensionless
surface mass density, $\kappa(\bmath{\theta})$, by 
\be \label{eq:psikappa}
\kappa(\bmath{\theta})=\frac{1}{2}\nabla^2\psi(\bmath{\theta}).
\ee
}

Strong lenses sit in our real Universe -- they are not truely isolated;
there is some mass along the line of sight coming from intervening structures.
These structures may be physically associated with the lens e.g. other members
of a group or cluster, or they may be the result of chance alignments at any redshift. 
Whilst it is rare for three galaxies to line up well enough for both of the background sources
to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, the shear scale of dark matter halos
makes a partial alignment -- with the additional structure acting as a perturbing weak lens -- 
a near certainty \citep{Vale+White2003,HilbertEtal2007}. These additional
structures induce an external shear which can be recovered in the mass-modelling
of the lens, but they can also introduce a convergence that
is undetectable from the image positions, shapes and relative fluxes; {\it the
mass sheet-degeneracy} \citep{FalcoEtal1985}.

If the intrinsic source brightness changes, the observed brightness in each image does not change
simultaneously - there is a time delay, $\Delta t$, between the brightness change in
each image. Despite the invariance of the image positions and fluxes, under the addition
of an external convergence (and appropriate re-scaling of the lens potential) the relative Fermat potential between the images is {\it not}
invariant, and so the time delay changes. As such it is necessary to include $\kappa$
in the lens modelling, if cosmological parameters are to be estimated accurately and
precisely from observed time delays \citep{SuyuEtal2010}. If there is any external convergence, that is not included in the lens
modeling, then the time delay distance -- inferred assuming $\kappa = 0$ -- will be
$(1-\kappa)$ less than the true value of $\Dt$.
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=(1-\kappa) \Dt^{{\kappax = 0}}.
\ee
We can overcome this degeneracy with additional knowledge of $\kappa$.
Since $\kappa$ is typically small the absolute uncertainty on the estimate
of $\kappa$ correspends to the percentage uncertainty with which time-delay
distances can be infered.

For mass-sheets that are physically associated with the lens galaxy, the mass sheet
 affects the stellar dynamics of the
lens galaxy, so dynamical observations can break the internal
mass-sheet degeneracy by providing an additional estimate of the lens galaxy's mass
\citep[e.g.,][]{SuyuEtal2010}. In this work our aim is to quantify the lensing effect of mass-sheets that are not physically associated with the lens galaxy -- mass-sheets
lying infront or behind the lens -- and the uncertainties induced by reconstructing
those mass-sheets using simple prescriptions applied to potential astronomical observables.

@@

\section{Estimating external convergence: The halo model approximation}
\label{sec:model}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

For real lines of sight, we would like to estimate the external convergence given all
data available. Typically this data is in the form of a photometric galaxy
catalog, containing positions, brightnesses, colours, and possibly spectra for each galaxy in the
field of the lens, down to some magnitude
limit. Since we can not observe the surface mass density of each halo
directy, we need some way of estimating
the mass of each halo in the catalog, so that we can compute its
contribution to the total $\kappax$ along the line of sight. This mass assignment recipe will be uncertain, but
can be calibrated with cosmological simulations.

Transforming the photometometric galaxy catalogue into a  catalogue of
halo masses, positions, and redshifts, will enable us to attempt a
reconstruction of the convergence induced by every halo near a line of sight.
This is the subject of Section \ref{subsec:halos}. We also need to account for
the mass in the Universe that is not associated with the galaxies we can see
-- this is the subject of Section \ref{subsec:voids}.

Cosmological dark matter simulations have shown that dark matter
halos are well approximated by NFW profiles \citep{NFW}, with a density
profile given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}},
\ee
where $r_{s}$ is a characteristic radius of the cluster, representing where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is related
to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $r_{200}$ is the radius at 
which 
\be
\rho(r_{200}) = 200\rho_{{\mathrm{crit}}}(z), \text{ where } \rho_{\mathrm{crit}}(z) \equiv \frac{3H^{2}(z)}{8\pi G}
\ee
and $c$ is the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less concentrated,
but there is some scatter; we use the relation of \citet{Neto2007} to estimate $c$
from the mass enclosed within $r_{200}$, which we denote as $M_{200}$. We find our results do not change if the \citet{MaccioEtal2008} Mass--concencetration relation is used instead. The best fit relation of \citet{Neto2007} is given by
\be
c_{200} = 4.67 (M_{200}/10^{14} h^{-1} \Msun)^{-0.11}.
\ee
$\rho_0$ is calculated from the critical density and concentration parameter:
\be
\rho_0 = \rho_{\mathrm{crit}}(z)\frac{200}{3}\frac{c^{3}}{\ln(1+c)-c/(1+c)}.
\ee
